### 2021-03-15 manuscript/Results.tex

\subsection{"Real-time phase": IMN}

\subsubsection{Development of ecPoint-based products}

\begin{figure}
\centerline{\includegraphics[width=19pc]{manuscript/Figures/Results_ecPoint_Product_CR.png}}
\caption{Product based on ecPoint-Rainfall created at IMN to be used by forecasters during the "real-time phase". The product consists in displaying the ecPoint-Rainfall's 85th percentile over Costa Rica.}
\label{ecPoint_Product_CR}
\end{figure}

\subsubsection{Subjective Verification}

\begin{figure}
\centerline{\includegraphics[width=19pc]{manuscript/Figures/Results_Obs_CaseStudy_CR.png}}
\caption{Panel (a) shows the climate regions of Costa Rica defined by IMN. Panels from (a) to (f) show 12-hourly rainfall observations during the extreme rainfall event occurred between October 3rd and 5th, 2018.}
\label{Obs_CaseStudy_CR}
\end{figure}

\begin{figure*}
\centerline{\includegraphics[width=39pc]{manuscript/Figures/Results_CaseStudy_Comparison_Models_CR.png}}
\caption{Forecast evolution (from day 7 to day 1) for the rainfall event occurred between October 4th at 12am and October 5th at 0am (see Fig.\ref{Obs_CaseStudy_CR}d to compare the forecasts with the observations). From the left, the first column shows the 85th percentile for ecPoint-Rainfall (as used operationally by IMN), the second column shows the 85th percentile for the raw ENS, the third column shows the 99th percentile for ecPoint-Rainfall, the fourth column shows the wettest member of the raw ENS, and the fifth column shows the deterministic forecast from WRF-1.5 (spatial resolution of 1.5 km) typically used by IMN in the forecasts of extreme localized rainfall events. The shown raw ECMWF ENS and ecPoint-Rainfall forecasts correspond to runs at 12 UTC, which is the first available run from Europe to IMN forecasters due to time difference between Europe and America (UTC-6). The displayed WRF-1.5 forecasts correspond to runs at 18 UTC, which is the first run available to IMN forecasters in the morning for their in-house models. The colour scheme of the second column plots have been modified, compared to the original IMN products (see Fig.\ref{ecPoint_Product_CR}d), to standardize all the plots in this figure, and make easier the forecasts comparison between models.}
\label{Results_CaseStudy_Comparison_Models_CR}
\end{figure*}

\subsection{"Real-time phase": OMSZ}
\subsubsection{Development of ecPoint-based products}
\subsubsection{Subjective Verification}

\subsection{"Offline phase": IMN}

\begin{figure}
\centerline{\includegraphics[width=19pc]{manuscript/Figures/Results_CaseStudy_Comparison_Prob_Perc.png}}
\caption{Panels (a) and (c) show the probabilities of not exceeding 50 mm/12h, and panels (b) and (d) show the 99th percentile, both for ecPoint-Rainfall forecasts. Panels (a) and (b) correspond to the forecast on 2018/10/04 at 00 UTC (t+18,t+30), which correspond to the period between 2018/10/04 12am and 201/10/05 0am (local time). Panels (c) and (d) correspond to the forecast on 2018/10/04 at 12 UTC (t+18,t+30), which correspond to the period between 2018/10/05 0am and 201/10/05 12am (local time).} 
\label{CaseStudy_Comparison_Prob_Perc}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=19pc]{manuscript/Figures/Results_CaseStudy_CDFs_CR.png}}
\caption{CDFs for ecPoint-Rainfall (in blue) and raw ENS (in red). Panels (a) and (b) display the CDFs for a location representative of the south coast of the Nicoya peninsula (lat=9.82,lon=-84.94). Panels (c) and (d) display the CDFs for a location representative of the inland parts of the Nicoya peninsula (lat=10.08,lon=-85.47). Panel (a) and (c) correspond to the forecast on 2018/10/04 at 00 UTC (t+18,t+30), which correspond to the period between 2018/10/04 12am and 201/10/05 0am (local time). Panel (b) and (d) correspond to the forecast on 2018/10/04 at 12 UTC (t+18,t+30), which correspond to the period between 2018/10/05 0am and 201/10/05 12am (local time).}
\label{Results_CaseStudy_CDFs_CR}
\end{figure}

\subsection{"Offline phase": OMSZ}

\subsection{Revised ecPoint-Rainfall guidelines}

\begin{figure*}
\centerline{\includegraphics[width=39pc]{manuscript/Figures/Results_Correlation_Forecasters_ecPoint.png}}
\caption{Comparison between the forecasters approach when building their mental model during the forecast creation process, and the ecPoint post-processing methodology.}
\label{Results_Correlation_Forecasters_ecPoint}
\end{figure*}